---
title: "Datathon 2020"
author: "Congwei Yang"
date: "2/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Reading data

```{r}
data1=read.csv(file="Part 1.csv", header=T)
```

```{r}
data2=read.csv(file="Part 2.csv", header=T)
```

```{r}
data3=read.csv(file="Part 3.csv", header=T)
```

```{r}
data4a=read.csv(file="Part 4a.csv", header=T)
```

```{r}
data4b=read.csv(file="Part 4b.csv", header=T)
```

```{r}
data5=read.csv(file="Part 5.csv", header=T)
```

bind data together

```{r}
data=rbind(data1, data2, data3, data4a, data4b, data5)
```

data cleasing and processing

```{r}
A=table(data$GEO.id2,data$NAICS.display.label)
```

```{r}
Ad=data.frame(A)
zipcode=Ad$Var1[which(Ad$Freq!=0)]
segment=as.data.frame(table(zipcode))
segment
```

```{r}
#A00200 salary and wage amount
#A04800 taxable income
#A06500 income tax amount
data2=read.csv(file="12zpallagi.csv", header=T)
data2$zipcode=as.factor(data2$zipcode)
```

```{r}
library(dplyr)
newdata = data2 %>% 
  group_by(zipcode) %>% 
  summarise(hahah = sum(A00200), lalal=sum(MARS4))
```

```{r}
segment$zipcode=as.numeric(as.character(segment$zipcode))
newdata$zipcode=as.numeric(as.character(newdata$zipcode))
```

```{r}
library(dplyr)
total = inner_join(segment, newdata, by="zipcode")
```

```{r}
salesdata=data
```

```{r}
levels(salesdata$RCPSZFE.id)=c(0,0,50000,175000,375000,750000,1000000,0,0,0,0)
salesdata=salesdata[-1,]
salesdata$RCPSZFE.id=as.numeric(as.character(salesdata$RCPSZFE.id))
salesdata$ESTAB=as.numeric(as.character(salesdata$ESTAB))
```

```{r}
salesdata2 = salesdata %>% 
  mutate(sales = RCPSZFE.id * ESTAB) %>% 
  group_by(GEO.id2) %>% 
  summarise(hahah = sum(sales))

salesdata2

colnames(salesdata2)=c('zipcode', 'salesamount')
salesdata2$zipcode=as.numeric(as.character(salesdata2$zipcode))
```

```{r}
total = inner_join(total, salesdata2, by="zipcode")
```

```{r}
colnames(total)=c('zipcode', 'segment number', 'total wage', 'household head number', 'total market sale')
```

```{r}
colnames(total)=c('zipcode','segment number', 'average wage', 'household head number', 'average market sale')
```

```{r}
total$'average wage'=total$'average wage'/total$`household head number`
```

```{r}
total$`average market sale`=total$`average market sale`/total$`household head number`
```

```{r}
data3=read.delim(file='EC1271Z1.dat', header=T, sep="|")
```

```{r}
newdata3 = data3 %>% 
  group_by(ZIPCODE) %>% 
  summarise(A = sum(ESTAB))
```

```{r}
colnames(newdata3)=c('zipcode', 'entertainment number')
total=left_join(total, newdata3, by='zipcode')
```

```{r}
total$`entertainment number`[is.na(total$`entertainment number`)]=0
```

```{r}
data4=read.delim(file="EC1272Z1.dat", header=T, sep='|')
```

```{r}
newdata4 = data4 %>% 
  group_by(ZIPCODE) %>% 
  summarise(A = sum(ESTAB))
```

```{r}
colnames(newdata4)=c('zipcode', 'F/A number')
total=left_join(total, newdata4, by='zipcode')
```

```{r}
colnames(total)=c('zipcode', 'segment_n', 'wage', 'household_number', 'mean_sale','entertainment_n','F/A_n')
total$`F/A_n`[is.na(total$`F/A_n`)]=0
```

```{r}
#house hold number = 0 should be discarded
zerohh=which(total$household_number==0)
```


```{r}
total=total[-zerohh, ]
```

Data spliting

```{r}
library(caret)
set.seed(47)
trainindex = createDataPartition(total$zipcode, p = 0.8,
                                 list = FALSE, times = 1)
total_trn = total[trainindex,]
total_tst = total[-trainindex,]
```

training models: 

random forest model

```{r}
fitcontrol = trainControl(method = "cv",
                          number = 5,
                          )
set.seed(47)
rfmod = train(
  mean_sale ~ .-zipcode,
  data = total_trn,
  method = "ranger",
  trControl = fitcontrol,
  metric = "RMSE"
)
#rfmod_rmse = 
rfmod$result
```

```{r}
#TEST RMSE
rfpred = predict(rfmod, newdata = total_tst)
rf_tst_rmse = postResample(pred = rfpred, obs = total_tst$mean_sale)
rf_tst_rmse
```

linear model

```{r}
set.seed(47)
lmmod = train(
  mean_sale ~ .-zipcode,
  data = total_trn,
  method = "lm",
  trControl = fitcontrol,
  metric = "RMSE"
)
#rfmod_rmse = 
lmmod$result
```

```{r}
#TEST RMSE
lmpred = predict(lmmod, newdata = total_tst)
lm_tst_rmse = postResample(pred = lmpred, obs = total_tst$mean_sale)
lm_tst_rmse
```

neural network model

```{r}
set.seed(47)
nnmod = train(
  mean_sale ~ .-zipcode,
  data = total_trn,
  method = "brnn",
  trControl = fitcontrol,
  metric = "RMSE"
)
#rfmod_rmse = 
nnmod$result
```

```{r}
#TEST RMSE
nnpred = predict(nnmod, newdata = total_tst)
nn_tst_rmse = postResample(pred = nnpred, obs = total_tst$mean_sale)
nn_tst_rmse
```

gbm model

```{r}
set.seed(47)
gbmmod = train(
  mean_sale ~ .-zipcode,
  data = total_trn,
  method = "gbm",
  trControl = fitcontrol,
  metric = "RMSE"
)
#rfmod_rmse = 
gbmmod$result
```

```{r}
#TEST RMSE
gbmpred = predict(gbmmod, newdata = total_tst)
gbm_tst_rmse = postResample(pred = gbmpred, obs = total_tst$mean_sale)
gbm_tst_rmse
```

model comparison

```{r}
#model comparison
result = data.frame(
  RMSE = c(rfmod$result$RMSE,lmmod$result$RMSE,nnmod$result$RMSE,gbmmod$result$RMSE),
  Model_Names = c(
    rep("Random Forest", nrow(rfmod$result)),
    "Linear Regression",
    rep("Neural Network", nrow(nnmod$result)),
    rep("Gbm", nrow(gbmmod$result))
  )
)

ggplot(result, aes(
  x = Model_Names,
  y = RMSE,
  fill = Model_Names,
  color = factor(Model_Names)
)) +
  scale_y_continuous(name = "RMSE")+
  geom_boxplot(alpha = 0.3)+
  theme(panel.background = element_blank(),legend.key = element_rect(fill = "white"), axis.ticks.x = element_blank())
```

making predicted value

```{r}
prediction=predict(rfmod, newdata=total)
total=cbind(total, as.data.frame(prediction))
```

```{r}
total = total %>% 
  mutate(sale_difference = prediction - mean_sale)
```

create rating and bind it to the data

```{r}
rating=total$sale_difference/100000
summary(rating)
```

```{r}
rated=cbind(total, as.data.frame(rating))
```

```{r}
rankingdata=rated
```

```{r}
rankingdata=rankingdata[order(rankingdata$rating),]
```

```{r}
ranking=rev(c(1:length(rankingdata$rating)))
rankingdata=cbind(rankingdata, as.data.frame(ranking))
```

read and merge position data for later visualization

```{r}
codata=read.csv(file="co.csv", header=T, sep=";")
```

```{r}
codata=codata %>%
  select(Zip, Latitude, Longitude)
colnames(codata)=c("zipcode", "Latitude", "Longitude")
```

```{r}
completedata=left_join(rankingdata, codata, by="zipcode")
```

Visualization: 

```{r}
d1=data[which(data$RCPSZFE.id=='001'),]
d2=data[which(data$RCPSZFE.id=='1'),]
```

```{r}
pdata=rbind(d1,d2)
```

```{r}
pdata$GEO.id2=as.numeric(as.character(pdata$GEO.id2))
```

Shiny using zipcode

```{r}
library(googleway)

## Not run: 
library(shiny)
library(googleway)

ui <- fluidPage(
  textInput(inputId = 'zipcode', 
            label = 'Enter Zip code here', 
            ), 
  submitButton("Click me"), 
  google_mapOutput("map"), 
  dataTableOutput("info"))

server <- function(input, output, session) {
  set_key("AIzaSyDk0GcRhiZ_NGyeYrmBrDiahnoJrJlu34Q")
  

  output$map <- renderGoogle_map({
    posdata=completedata %>% select(zipcode, Latitude, Longitude)
    google_map(location = c(posdata$Latitude[which(posdata$zipcode==input$zipcode)], posdata$Longitude[which(posdata$zipcode==input$zipcode)]),
                zoom = 14)
  })
  infodata=completedata %>% select(zipcode, ranking, rating,segment_n)
  output$info <- renderDataTable({
    infodata[which(infodata$zipcode==input$zipcode),]
  })
  
  
  
}

shinyApp(ui, server)
```

Shiny using ranking

```{r}
library(googleway)

## Not run: 
library(shiny)
library(googleway)

ui <- fluidPage(
  textInput(inputId = 'ranking', 
            label = 'Enter ranking code here', 
            ), 
  submitButton("Click me"), 
  google_mapOutput("map"), 
  dataTableOutput("info"))

server <- function(input, output, session) {
  set_key("AIzaSyDk0GcRhiZ_NGyeYrmBrDiahnoJrJlu34Q")
  

  output$map <- renderGoogle_map({
    posdata=completedata %>% select(zipcode, Latitude, Longitude, ranking)
    google_map(location = c(posdata$Latitude[which(posdata$ranking==input$ranking)], posdata$Longitude[which(posdata$ranking==input$ranking)]),
                zoom = 14)
  })
  infodata=completedata %>% select(zipcode, ranking, rating,segment_n)
  output$info <- renderDataTable({
    infodata[which(infodata$ranking==input$ranking),]
  })
}

shinyApp(ui, server)
```

a few ggplot to explore other characteristics

```{r}
ggplot(completedata, mapping=aes(household_number, rating))+
         geom_point(size=1, col='deepskyblue')
  
```

```{r}
ggplot(completedata, mapping=aes(ranking, rating))+
         geom_point(size=1, col='deepskyblue')
```

```{r}
ggplot(completedata, mapping=aes(rating, wage))+
         geom_point(size=1, col='deepskyblue')
```


